Introducing ICBe: An Event Extraction Dataset from Narratives about International Crises

### 2023-08-20 replication_data/rmd/02_format_and_clean.rmd
v1.1 first self contained hopefully
R&R

dictionary_actions  <- readxl::read_excel(target_file, sheet="actions") %>% dplyr::distinct() 
dictionary_actors    <- readxl::read_excel(target_file, sheet="actors") %>% dplyr::select(crisno,value, value_qcode=value_normalized_wikidata_id, value_normalized=value_normalized_label) %>% dplyr::distinct() %>%
                        full_join(varname_count %>% dplyr::select(varname) %>% filter(varname %>% str_detect("actor")),  by = character())

@@

normalization_dictionaries <- bind_rows(dictionary_actions %>% rename(varname=variable, varname_normalized=variable_normalized),
                                        dictionary_locations,
                                        dictionary_condition_locations,
                                        dictionary_units,
                                        dictionary_domains,
                                        dictionary_counts
                                        )

@@

icb_long_spans_clean_normalized %>% saveRDS(paste0(here::here(), "/replication_data/temp/ICBe_V1.1_long_clean.Rds"))

>>>>

dictionary_actions  <- readxl::read_excel(target_file, sheet="actions") %>% dplyr::distinct() 
dictionary_condition_actions  <- readxl::read_excel(target_file, sheet="condition_actions") %>% dplyr::distinct()

dictionary_actors    <- readxl::read_excel(target_file, sheet="actors") %>% dplyr::select(crisno,value, value_qcode=value_normalized_wikidata_id, value_normalized=value_normalized_label) %>% dplyr::distinct() %>%
                        full_join(varname_count %>% dplyr::select(varname) %>% filter(varname %>% str_detect("actor")),  by = character())

@@

normalization_dictionaries <- bind_rows(dictionary_actions %>% rename(varname=variable, varname_normalized=variable_normalized),
                                        dictionary_condition_actions %>% rename(varname=variable, varname_normalized=variable_normalized),
                                        dictionary_locations,
                                        dictionary_condition_locations,
                                        dictionary_units,
                                        dictionary_domains,
                                        dictionary_counts
                                        )

@@

icb_long_spans_clean_normalized %>% saveRDS(paste0(here::here(), "/replication_data/out/ICBe_V1.1_long_clean.Rds"))

### 2023-08-20 replication_paper/psrm_draft/ICBe_manuscript.Rmd
R&R

icb_actors <- read.csv(paste0(here::here(), "/replication_data/in/icb2v14.csv"))
```

@@

The ICBe ontology follows a hierarchical design philosophy where a smaller number of significant decisions are made early on and then progressively refined into more specific details [@brustIntegratingDomainKnowledge2020].[^5] Each coder was instructed to first thoroughly read the full crisis narrative and then presented with a custom graphical user interface (GUI) (SI Appendix 2.1). Coders then proceeded sentence by sentence, choosing the number of events (0-3) that occurred, the highest behavior (thought, speech, or activity), a set of players, whether the means were primarily armed or unarmed, whether there was an increase or decrease in aggression (uncooperative/escalating or cooperative/de-escalating), and finally one or more specific and non-mutually exclusive activities. Some additional details were always collected (e.g. location and timing) while other details were only collected if appropriate (e.g. force size, fatalities, domains, units). While each event was matched to a sentence, coders could fill in details outside that sentence (e.g. antecedents to pronouns). We reviewed, standardized, and normalized where coders listed a behavior, actor, or location outside the ontology.[^6] We find `r codings_long_agreement %>% filter(varname_normalized %>% str_detect("actor")) %>% pull(value_normalized) %>% unique() %>% length()` actors and 117 different behaviors.

@@

Each crisis was typically assigned to 2 expert coders and 2 novice coders with an additional tie-breaking expert coder assigned to sentences with high disagreement.[^8] For the purposes of measuring intercoder agreement and consensus, we temporarily disaggregate the unit of analysis to the Coder-Crisis-Sentence-Tag (n=`r icb_long_clean %>% nrow() %>% scales::comma()`), where a tag is any unique piece of information a coder can associate with a sentence such as an actor, date, behavior, etc. We then aggregate those tags into final events (n=`r ICBe_events_agreed %>% nrow() %>% scales::comma()`), using a consensus procedure (SI Appendix 2.2) that requires a tag to have been chosen by at least one expert coder and either a majority of expert or novice coders. This screens noisy tags that no expert considered possible but leverages novice knowledge to tie-break between equally plausible tags chosen by experts. Requiring sentence-tag matching may underestimate agreement but minimizes the inclusion of noise and allows for additional validation.

>>>>

icb_actors <- read.csv(paste0(here::here(), "/replication_data/in/icb2v14.csv"))

@@

The ICBe ontology follows a hierarchical design philosophy where a smaller number of significant decisions are made early on and then progressively refined into more specific details [@brustIntegratingDomainKnowledge2020].[^5] Each coder was instructed to first thoroughly read the full crisis narrative and then presented with a custom graphical user interface (GUI) (SI Appendix 2.1). Coders then proceeded sentence by sentence, choosing the number of events (0-3) that occurred, the highest behavior (thought, speech, or action), a set of players, whether the means were primarily armed or unarmed, whether there was an increase or decrease in aggression (uncooperative/escalating or cooperative/de-escalating), and finally one or more specific and non-mutually exclusive activities. Some additional details were always collected (e.g. location and timing) while other details were only collected if appropriate (e.g. force size, fatalities, domains, units). While each event was matched to a sentence, coders could fill in details outside that sentence (e.g. antecedents to pronouns). We reviewed, standardized, and normalized where coders listed a behavior, actor, or location outside the ontology.[^6]

@@

Each crisis was typically assigned to 2 expert coders and 2 novice coders with an additional tie-breaking expert coder assigned to sentences with high disagreement.[^8] For the purposes of measuring intercoder agreement and consensus, we temporarily disaggregate the unit of analysis to the Coder-Crisis-Sentence-Tag (n=`r icb_long_clean %>% nrow() %>% scales::comma()`), where a tag is any unique piece of information a coder can associate with a sentence such as an actor, date, behavior, etc. We then aggregate those tags into final events (n=`r ICBe_events_agreed %>% nrow() %>% scales::comma()`), using a consensus procedure (SI Appendix 2.2) that requires a tag to have been chosen by at least one expert coder and either a majority of expert or novice coders. This screens noisy tags that no expert considered possible but leverages novice knowledge to tie-break between equally plausible tags chosen by experts. Requiring sentence-tag matching may underestimate agreement but minimizes the inclusion of noise and allows for additional validation. Once filtered for agreement, we find `r codings_long_agreement %>% filter(varname_normalized %>% str_detect("actor")) %>% pull(value_normalized) %>% unique() %>% length()` actors and `r codings_long_agreement %>% filter(varname_normalized %>% str_detect("act_cooperative|interact_increasecoop|act_uncooperative|interact_decreasecoop|act_escalate|interact_escalate|act_deescalate|interact_deescalate|sayintkind|thinkkind")) %>% filter(grepl("[A-Za-z]", value_normalized)) %>% pull(value_normalized) %>% unique() %>% length()` different behaviors: `r codings_long_agreement %>% filter(varname_normalized %>% str_detect("thinkkind")) %>% filter(grepl("[A-Za-z]", value_normalized)) %>% pull(value_normalized) %>% unique()  %>% length()` thought, `r codings_long_agreement %>% filter(varname_normalized %>% str_detect("sayintkind")) %>% filter(grepl("[A-Za-z]", value_normalized)) %>% pull(value_normalized) %>% unique()  %>% length()` speech, and `r codings_long_agreement %>% filter(varname_normalized %>% str_detect("act_cooperative|interact_increasecoop|act_uncooperative|interact_decreasecoop|act_escalate|interact_escalate|act_deescalate|interact_deescalate")) %>% pull(value_normalized) %>% unique() %>% length()` actions.

### 2023-08-20 replication_paper/psrm_draft/ICBe_manuscript.tex
R&R

The ICBe ontology follows a hierarchical design philosophy where a
smaller number of significant decisions are made early on and then
progressively refined into more specific details (Brust and Denzler
2020).\footnote{This process quickly focuses the coder on a smaller
  number of relevant options while also allowing them to apply multiple
  tags if the sentence explicitly includes more than one or there is
  insufficient evidence to choose only one tag. The guided coding
  process also allows for the possibility that earlier coarse decisions
  have less error than later fine-grained decisions.} Each coder was
instructed to first thoroughly read the full crisis narrative and then
presented with a custom graphical user interface (GUI) (SI Appendix
2.1). Coders then proceeded sentence by sentence, choosing the number of
events (0-3) that occurred, the highest behavior (thought, speech, or
activity), a set of players, whether the means were primarily armed or
unarmed, whether there was an increase or decrease in aggression
(uncooperative/escalating or cooperative/de-escalating), and finally one
or more specific and non-mutually exclusive activities. Some additional
details were always collected (e.g.~location and timing) while other
details were only collected if appropriate (e.g.~force size, fatalities,
domains, units). While each event was matched to a sentence, coders
could fill in details outside that sentence (e.g.~antecedents to
pronouns). We reviewed, standardized, and normalized where coders listed
a behavior, actor, or location outside the ontology.\footnote{See the
  full codebook on Github Repository
  \href{https://urldefense.com/v3/__https://github.com/CenterForPeaceAndSecurityStudies/ICBEventData__;!!Mih3wA!WxDJtEczKfxGTh0S2Krunap8ReymFEL5iTWaSfOHeqlSdyfRx77zmjBSWO1OAm13$}{ICBEventData}.}
We find 472 actors and 117 different behaviors.

@@

Each crisis was typically assigned to 2 expert coders and 2 novice
coders with an additional tie-breaking expert coder assigned to
sentences with high disagreement.\footnote{Expert coders were graduate
  students or postgraduates who collaboratively developed the ontology
  and documentation for the codebook. Undergraduate coders were students
  who engaged in classroom workshops.} For the purposes of measuring
intercoder agreement and consensus, we temporarily disaggregate the unit
of analysis to the Coder-Crisis-Sentence-Tag (n=993,740), where a tag is
any unique piece of information a coder can associate with a sentence
such as an actor, date, behavior, etc. We then aggregate those tags into
final events (n=18,783), using a consensus procedure (SI Appendix 2.2)
that requires a tag to have been chosen by at least one expert coder and
either a majority of expert or novice coders. This screens noisy tags
that no expert considered possible but leverages novice knowledge to
tie-break between equally plausible tags chosen by experts. Requiring
sentence-tag matching may underestimate agreement but minimizes the
inclusion of noise and allows for additional validation.

>>>>

The ICBe ontology follows a hierarchical design philosophy where a
smaller number of significant decisions are made early on and then
progressively refined into more specific details (Brust and Denzler
2020).\footnote{This process quickly focuses the coder on a smaller
  number of relevant options while also allowing them to apply multiple
  tags if the sentence explicitly includes more than one or there is
  insufficient evidence to choose only one tag. The guided coding
  process also allows for the possibility that earlier coarse decisions
  have less error than later fine-grained decisions.} Each coder was
instructed to first thoroughly read the full crisis narrative and then
presented with a custom graphical user interface (GUI) (SI Appendix
2.1). Coders then proceeded sentence by sentence, choosing the number of
events (0-3) that occurred, the highest behavior (thought, speech, or
action), a set of players, whether the means were primarily armed or
unarmed, whether there was an increase or decrease in aggression
(uncooperative/escalating or cooperative/de-escalating), and finally one
or more specific and non-mutually exclusive activities. Some additional
details were always collected (e.g.~location and timing) while other
details were only collected if appropriate (e.g.~force size, fatalities,
domains, units). While each event was matched to a sentence, coders
could fill in details outside that sentence (e.g.~antecedents to
pronouns). We reviewed, standardized, and normalized where coders listed
a behavior, actor, or location outside the ontology.\footnote{See the
  full codebook on Github Repository
  \href{https://urldefense.com/v3/__https://github.com/CenterForPeaceAndSecurityStudies/ICBEventData__;!!Mih3wA!WxDJtEczKfxGTh0S2Krunap8ReymFEL5iTWaSfOHeqlSdyfRx77zmjBSWO1OAm13$}{ICBEventData}.}

@@

Each crisis was typically assigned to 2 expert coders and 2 novice
coders with an additional tie-breaking expert coder assigned to
sentences with high disagreement.\footnote{Expert coders were graduate
  students or postgraduates who collaboratively developed the ontology
  and documentation for the codebook. Undergraduate coders were students
  who engaged in classroom workshops.} For the purposes of measuring
intercoder agreement and consensus, we temporarily disaggregate the unit
of analysis to the Coder-Crisis-Sentence-Tag (n=993,731), where a tag is
any unique piece of information a coder can associate with a sentence
such as an actor, date, behavior, etc. We then aggregate those tags into
final events (n=18,783), using a consensus procedure (SI Appendix 2.2)
that requires a tag to have been chosen by at least one expert coder and
either a majority of expert or novice coders. This screens noisy tags
that no expert considered possible but leverages novice knowledge to
tie-break between equally plausible tags chosen by experts. Requiring
sentence-tag matching may underestimate agreement but minimizes the
inclusion of noise and allows for additional validation. Once filtered
for agreement, we find 472 actors and 119 different behaviors: 12
thought, 13 speech, and 94 actions.
